---
editor_options: 
  markdown: 
    wrap: sentence
---

# Data Wrangling {#tidy}

## Overview (Mexico Toy Sales project)

Sales & inventory data for a fictitious chain of toy stores in Mexico called Maven Toys, including information about products, stores, daily sales transactions, and current inventory levels at each location.

In this project, you'll work with a dataset with 830,940 records and 14 fields for sales & inventory data for a fictitious chain of toy stores in Mexico called Maven Toys, including information about products, stores, daily sales transactions, and current inventory levels at each location pulled from ![Maven Analytics](https://maven-datasets.s3.amazonaws.com/Maven+Toys/Maven+Toys+Data.zip)

Some of the things you'll learn in this project are:

-   How to organize information using basic Python data structures.

-   How to import data from CSV files and clean it using the pandas library.

-   How to create data visualizations like scatter and box plots.

-   How to examine the relationship between two variables using correlation.

### Recommended Analysis

1.  Which product categories drive the biggest profits?

2.  Is this the same across store locations?

3.  Can you find any seasonal trends or patterns in the sales data?

4.  Are sales being lost with out-of-stock products at certain locations?

How much money is tied up in inventory at the toy stores?
How long will it last?

## Import data

There are so many packages in **R** that can be used to extract data into Rstudio environment.
For **Mexico Toy Sales** data, we will use `tidyverse` package.

```{r, warning=FALSE, message=FALSE}
library(tidyverse) # core for importing data
library(fs) # handling files
temp <- tempfile()
download.file("https://maven-datasets.s3.amazonaws.com/Maven+Toys/Maven+Toys+Data.zip",temp, model="wb")

## create a directory
fs::dir_create("data/raw_data")

# reading files from the unzipped folder
inventory <- read_csv(unz(temp, "inventory.csv"))
stores <- read_csv(unz(temp, "stores.csv"))
products <- read_csv(unz(temp, "products.csv"))
sales <- read_csv(unz(temp, "sales.csv"))

# save files to local machine
inventory%>%write_csv("data/raw_data/inventory.csv")
stores%>%write_csv("data/raw_data/stores.csv")
products%>%write_csv("data/raw_data/products.csv")
sales%>%write_csv("data/raw_data/sales.csv")
```

### Data dictionary

| Field            | Description                                            |
|-------------------|-----------------------------------------------------|
| Store_ID         | Store ID                                               |
| Product_ID       | Product ID                                             |
| Stock_On_Hand    | Stock quantity of the product in the store (inventory) |
| Store_Name       | Store name                                             |
| Store_City       | City in Mexico where the store is located              |
| Store_Location   | Location in the city where the store is located        |
| Store_Open_Date  | Date when the store was opened                         |
| Sale_ID          | Sale ID                                                |
| Date             | Date of the transaction                                |
| Units            | Units sold                                             |
| Product_ID       | Product ID                                             |
| Product_Name     | Product name                                           |
| Product_Category | Product Category                                       |
| Product_Cost     | Product cost (\$USD)                                   |
| Product_Price    | Product retail price (\$USD)                           |

Now we need to join our data from different sources into one table.
We will use `left_join` to achieve that.
Lets read the data from our local machine using `read_csv` first before we join.

```{r, message=FALSE}
inventory <- read_csv("data/raw_data/inventory.csv")
stores <- read_csv("data/raw_data/stores.csv")
products <- read_csv("data/raw_data/products.csv")
sales <- read_csv("data/raw_data/sales.csv")
```

### Data Understanding

Before we join the 4 datasets, we need to explore all and gain a bigger picture of the information presented in these datasets.

#### Stores Summary

Lets check the number of stores from this dataset:

```{r}
numOfCities <- length(unique(stores$Store_City))
numOfCities
```

Lets check the number of cities in which these stores are located from `stores` dataset:

```{r}
numOfStores <- length(unique(stores$Store_ID))
numOfStores
```

Lets check the city location in which these stores are situated from `stores` dataset:

```{r}
locType <- unique(stores$Store_Location)

print(paste("There are", length(locType), "locations in which these stores are located. These locations are namely ",locType[1],",",locType[2],",",locType[3], "and", locType[4]))
```

We need to check the newest and oldest stores from `stores` dataset:

```{r}
stores%>%
  mutate(Year_Built=year(Store_Open_Date))%>%
  mutate(Store_Age = 2022-Year_Built)%>%
  arrange(desc(Store_Age))%>%
  select(Store_Name, Store_City, Store_Location, Year_Built, Store_Age)%>%
  tail(1)%>%
  kableExtra::kable(format = "html", caption = "THE NEWEST STORE")
```

```{r}
stores%>%
  mutate(Year_Built=year(Store_Open_Date))%>%
  mutate(Store_Age = 2022-Year_Built)%>%
  arrange(desc(Store_Age))%>%
  select(Store_Name, Store_City, Store_Location, Year_Built, Store_Age)%>%
  head(1)%>%
  kableExtra::kable(format = "html", caption = "THE OLDEST STORE")

```
#### Products Summary

```{r}
products <- products%>%
  # Renaming Columns: set_names()
  set_names(names(.)%>%tolower())


products_tbl <- products%>%
  
  # remove $ sign and convert product cost and price to numeric
  mutate(product_cost = as.numeric(gsub('[\\$,]', '', product_cost)),
         product_price = as.numeric(gsub('[\\$,]', '', product_price)),
         profit = product_price - product_cost,
         percent_profit = round(profit/product_cost*100,3))

# summarize total num of products per category
products_tbl%>%
  group_by(product_category)%>%
  summarise(count = n())%>%
  ungroup()

# tabulate category with counts
table(products_tbl$product_category)
```
#### Inventory Summary

#### Sales Summary

#### Joining datasets

```{r}
toys_sales_joined_tbl <- inventory%>%
  
  # join inventroy with products datasets
  left_join(products, by=c("Product_ID"="Product_ID"))%>%
  
  # add stores dataset
  left_join(stores, by=c("Store_ID"="Store_ID"))%>%
  
  # add sales dataset
  left_join(sales, by=c("Store_ID"="Store_ID", "Product_ID"="Product_ID"))%>%
  
  # create year_built column for each store
  mutate(year_built = year(Store_Open_Date), store_age =  2022 - year(Store_Open_Date))%>%
  
  # Renaming Columns: set_names()
  set_names(names(.)%>%tolower())%>%
  
  # view joined dataset
  glimpse()

```
```{r}
toys_sales_joined_tbl %>% 
  select(store_name, store_city, store_location, year_built, store_age)%>%
  arrange(desc(store_age))%>%
  head(1)%>%
  kableExtra::kable(format = "html", caption = "THE OLDEST STORE")
```

```{r}
toys_sales_joined_tbl %>% 
  select(store_name, store_city, store_location, year_built, store_age)%>%
  arrange(desc(store_age))%>%
  tail(1)%>%
  kableExtra::kable(format = "html", caption = "THE NEWEST STORE")
```

```{r}
```
